Evaluating Cultural Transmission in Bronze Age burial rites of Central, Northern and North-western Europe using radiocarbon data

### 2019-03-31 article/02_Materials_and_Methods.Rmd
incorporating new results

Defining a measure of spatial distance for the above established main units of observation, the macro-regions, is only possible in a highly simplified way. A continuous distance metric such as least-cost path distance cannot be employed on this scale, but it is possible to distinguish ordinal distance classes. For the eight regions, four different categories of spatial distance were defined (Figure \ref{map_regions} B). The test for correlation between spatial and cultural distance works similar to the one between the two cultural distances above, except that the Mantel test has to be applied with Spearman's rank correlation coefficient [@spearmanProofMeasurementAssociation1904], because of the different data levels of the measurements (Table \ref{tab:dacta} AB).

@@

\begin{figure}
\centering
\includegraphics[width=\columnwidth]{correlation_time_series.jpeg}
\caption{\label{correlationtimeseries}Development of correlation between cultural and spatial distance. The results for the burial customs data (Table \ref{tab:dacta} AB) are displayed with a solid (\textit{burial type}) and dashed (\textit{burial construction}) line. The horizontal line marks the correlation coefficient of zero, where neither positive nor negative correlation can be assumed.}
\end{figure}

>>>>

Defining a measure of spatial distance for the above established main units of observation, the macro-regions, is only possible in a highly simplified way. A continuous distance metric such as least-cost path distance cannot be employed on this scale, but it is possible to distinguish ordinal distance classes. For the eight regions, four different categories of spatial distance were defined (Figure \ref{map_regions} B). The test for correlation between spatial and cultural distance works similar to the one between the two cultural distances above, except that the Mantel test has to be applied with Spearman's rank correlation coefficient [@spearmanProofMeasurementAssociation1904], because of the different data levels of the measurements (Table \ref{tab:dacta}).

@@

\begin{figure}
\centering
\includegraphics[width=\columnwidth]{correlation_time_series.jpeg}
\caption{\label{correlationtimeseries}Development of correlation between different measures of cultural and spatial distance. The results for the correlation of cultural and spatial distance (Table \ref{tab:dacta}) are displayed with a solid (\textit{burial type}) and a dashed (\textit{burial construction}) line. The results for the correlation of the two cultural distances (Table \ref{tab:dactb}) are displayed with a strong dotted line \textit{without} the correction for spatial distance (only here: Pearson correlation coefficient), and with a weak dotted line for the result \textit{with} the aforementioned correction. The horizontal line marks the correlation coefficient of zero, where neither positive nor negative correlation can be assumed.}
\end{figure}

### 2019-03-31 article/03_Results.Rmd
started to incorporate the new arrangements of Figure 6 and Partial Mantel test into the text
incorporating new results

Table \ref{tab:dacta} C shows that the distance networks of *burial type* and *burial construction* are not equivalent. There is also no significant negative correlation. The degree of agreement between two regions in the question of inhumation and cremation is therefore not suitable for predicting the similarity of developments in these regions concerning the dominance of flat or mound graves. Only for the time slot at the transition to the Middle Bronze Age from 1800 to 1600 calBC the 5% significance level of the Mantle test is almost undercut, thus suggesting a statistically significant relationship.

For the relationship between the SED networks and the spatial distance network several observations result from Table \ref{tab:dacta} AB: Cultural distance generally does not seem to correlate with spatial distance. For the variable *burial type* an exception with significant correlation between the distance measures only reveals itself for the Early Bronze Age time slot 2000 to 1800 calBC. In the Early Bronze Age, long before the universal switch to the Urnfield burial tradition, large regions directly adjacent to each other tend to see the same developments. In the Middle Bronze Age the SED developments of the *burial type* variable undergo the peaks already observed. The correlation between cultural and spatial distance during these peaks is low and even tilts between 1400 and 1200 calBC, at the transition to the Late Bronze Age, into the negative -- although not significantly. The cultural distance between several geographically distant regions is then smaller than between several directly adjacent regions. Looking at the SED network of *burial construction*, similar conclusions can be drawn: The distance values here are larger overall, and the SED differences within the individual spatial distance classes are clearly more accentuated. As well as in *burial type* the Early Bronze Age is by far the phase with the most substantial spatial correlation, though it is only significant for the time slot between 1800 and 1400 calBC. There is no time period with negative mean correlation for *burial construction* cultural distance and spatial distance, but also here the correlation coefficient approaches zero in the Late Bronze Age.

>>>>

Table \ref{tab:dactb} shows that the distance networks of *burial type* and *burial construction* are not equivalent. There is also no significant negative correlation. The degree of agreement between two regions in the question of inhumation and cremation is therefore not suitable for predicting the similarity of developments in these regions concerning the dominance of flat or mound graves. Only for the time slot at the transition to the Middle Bronze Age from 1800 to 1600 calBC the 5% significance level of the Mantle test is almost undercut, thus suggesting a statistically significant relationship. This moment of strong correlation is though not induced by similarity along the spatial dimension: The effect is still visible in the result of the Partial Mantel test, where the spatial distance is artificially counterbalanced.

For the relationship between the SED networks and the spatial distance network several observations result from Table \ref{tab:dacta}:

For the variable *burial type* an exception with significant correlation between the distance measures only reveals itself for the Early Bronze Age time slot 2000 to 1800 calBC. In the Early Bronze Age, long before the universal switch to the Urnfield burial tradition, large regions directly adjacent to each other tend to see the same developments. In the Middle Bronze Age the SED developments of the *burial type* variable undergo the peaks already observed. The correlation between cultural and spatial distance during these peaks is low and even tilts between 1600 and 1400 calBC into the negative -- although not significantly. The cultural distance between several geographically distant regions is then smaller than between several directly adjacent regions.

Looking at the SED network of *burial construction*, a slightly different conclusions can be drawn: The distance values here are larger overall, and the SED differences within the individual spatial distance classes are clearly more accentuated. As well as in *burial type* the Early Bronze Age is by far the phase with the most substantial spatial correlation, though it is only significant for the time slot between 2000 and 1400 calBC. There is no time period with negative mean correlation for *burial construction* cultural distance and spatial distance, but also here the correlation coefficient approaches zero in the Late Bronze Age.

In general cultural distance does not seem to correlate strongly with spatial distance, especially in the Middle and Late Bronze Age.
